---
title: "Kóvitinn"
output: flexdashboard::flex_dashboard
runtime: shiny
---

```{r global, include=FALSE}
from.web <- TRUE

# Check out:
# https://kieranhealy.org/blog/archives/2020/03/21/covid-19-tracking

library(lubridate)
library(highcharter)
library(tidyverse)
library(DT)
library(shiny)
source("R/functions.R")
# could possibly use wb stuff below:
countries.europa <- 
  read_rds("data/countries_europa.rds")
  # screws up the median
countries.europa <- countries.europa[countries.europa != "San Marino"]
if(from.web) {
  DATA <- 
    read_covid(variable = "confirmed") %>% 
    covid_tidy() %>%
    group_by(country, iso3c, date) %>% 
    summarise(n = sum(n)) %>% 
    ungroup() %>% 
    rename(confirmed = n) %>% 
    left_join(read_covid(variable = "deaths") %>% 
                covid_tidy() %>% 
                group_by(country, iso3c, date) %>%
                summarise(n = sum(n)) %>% 
                ungroup() %>% 
                rename(deaths = n)) %>% 
    mutate(deaths = case_when(country == "Iceland" &
                                date == ymd("2020-03-15") ~ 0,
                              TRUE ~ deaths)) %>% 
    left_join(read_rds("data/wbstats01.rds") %>% 
                select(iso3c, pop = population))
  # write_rds(DATA, "data/covid_2020-03-25.rds")
} else {
  DATA <- read_rds("data/covid_2020-03-25.rds")
}

DATAr <- 
  DATA %>%
  mutate(confirmed = confirmed / pop * 1e6,
         deaths = deaths / pop * 1e6) %>% 
  select(country, iso3c, date, confirmed, deaths) %>% 
  arrange(country, date) %>% 
  group_by(country, iso3c) %>% 
  mutate(day0c = date - min(date[confirmed >= 1]),
         day0d = date - min(date[deaths >= 1])) %>% 
  filter(day0c >= 0)

# NOTE: Median europe spooky on 1 day because of
#       San Marino
confirmed.europa.median <-
  DATAr %>%
  filter(country %in% countries.europa) %>%
  group_by(day0c) %>%
  summarise(confirmed = median(confirmed)) %>%
  ungroup()

res <- list()
for(i in 0L:20L) {
  print(i)
  res[[i + 1]] <-
    DATA %>%
    covid_doubling_time(i) %>%
    mutate(day = i)
}

confirmed.double <-
  bind_rows(res)

# creates an error
#res <- list()
#for(i in 0L:20L) {
#  print(i)
#  res[[i + 1]] <-
#    DATA %>%
#    covid_doubling_time(i, variable = deaths) %>%
#    mutate(day = i)
#}

deaths.double <-
  bind_rows(res)


countries <-
  c(countries.europa,   #Europe
    DATA %>% pull(country) %>% unique() %>% sort()) %>% 
  unique()



last.date <- DATA %>% pull(date) %>% max()
```

Sidebar {.sidebar data-width=175}
=====================================

```{r}
selectInput("country", "Select country:",
            choices = countries,
            selected = c("Iceland",
                         "Denmark",
                         "Estonia",
                         "Slovenia",
                         "Sweden",
                         "Norway",
                         "Greece",
                         "Italy",
                         "Finland",
                         "Slovakia"),
            multiple = TRUE)
radioButtons("log", "Y-scale", c("log", "ordinary"))

# Main path --------------------------------------------------------------------
output$pconfirmed <- renderPlot({ 
  p <-
    ggplot() +
    theme_bw(base_size = 14) +
    geom_line(data = confirmed.europa.median,
              aes(day0c, confirmed),
              lwd = 2,
              colour = "yellow") +
    geom_line(data = DATAr %>%
                filter(country %in% countries.europa) %>% 
                rename(group = country),
              aes(day0c, confirmed, group = group),
              colour = "grey", alpha = 0.5) +
    geom_line(data = DATAr %>% filter(country %in% input$country),
              aes(day0c, confirmed),
              colour = "darkgreen",
              lwd = 1) +
    geom_line(data = DATAr %>% 
                filter(country %in% input$country,
                       deaths >= 1),
              aes(day0c, deaths),
              colour = "red",
              lwd = 1) +
    facet_wrap(~ country) +
    scale_x_continuous(breaks = seq(0, 50, by = 5),
                       limits = c(0, 35)) +
    labs(x = "Days since more than 1 per million confirmed",
         y = NULL,
         caption = "Source: https://github.com/CSSEGISandData/COVID-19") +
    theme(legend.position = "none",
          panel.grid.minor = element_line(colour = "white"),
          panel.grid.major = element_line(size = 0.25))
  
  if(input$log == "log") {
    p + scale_y_log10()
  } else {
    p
  }
  
})

output$confirmed <- renderUI({ 
  plotOutput("pconfirmed")
})

# Confirmed - highchart --------------------------------------------------------

output$p_high_confirmed <- renderHighchart({
  h <-
    hchart(DATAr %>% 
           filter(country %in% input$country) %>% 
           mutate(confirmed = round(confirmed, 1)),
         "line", hcaes(x = day0c, y = confirmed, group = country)) %>%
    hc_add_theme(
      hc_theme_flatdark(
        chart = list(
          backgroundColor = "transparent",
          divBackgroundImage = "http://www.wired.com/images_blogs/underwire/2013/02/xwing-bg.gif"
        )
      )
    )
  
  if(input$log == "log") {
    h %>%  hc_yAxis(type = "logarithmic", 
             min = 1)
  } else {
    h
  }
  
  
    
})

output$high_confirmed <- renderUI({
  highchartOutput("p_high_confirmed")
})

# Death - highchart ------------------------------------------------------------
output$p_high_deaths <- renderHighchart({
  hchart(DATAr %>% 
           filter(country %in% input$country,
                  day0d >= 0) %>% 
           mutate(deaths = round(deaths, 1)),
         "line", hcaes(x = day0d, y = deaths, group = country)) %>%
    hc_add_theme(
      hc_theme_flatdark(
        chart = list(
          backgroundColor = "transparent",
          divBackgroundImage = "http://www.wired.com/images_blogs/underwire/2013/02/xwing-bg.gif"
        )
      )
    ) %>%
    hc_yAxis(type = "logarithmic",
             min = 1)
})

output$high_deaths <- renderUI({
  highchartOutput("p_high_deaths")
})
```

Most recent date: `r as.character(last.date)`

Confirmed
=====================================  

Column {.tabset}
-------------------------------------


### Confirmed

```{r}
uiOutput("confirmed")
```

### Confirmed - Single graph

```{r}
uiOutput("high_confirmed")
```

### Doubling time

```{r}
renderPlot({
  confirmed.double %>% 
    filter(country %in% input$country) %>% 
    ggplot(aes(-day, z)) +
    theme_bw(base_size = 14) +
    geom_point() +
    geom_smooth() +
    expand_limits(y = 1) +
    facet_wrap(~ country) +
    labs(title = "Doubling time in days of confirmed cases",
         y = "[days]",
         x = "Days since present day")
})
```

### Proportion confirmed

```{r}
renderPlot({
  DATA %>% 
    filter(date == max(date),
           country %in% countries.europa) %>% 
    mutate(p = confirmed / pop) %>% 
    ggplot(aes(reorder(country, p), p)) +
    theme_bw(base_size = 14) +
    geom_pointrange(aes(ymin = 0, ymax = p), lwd = 1) +
    geom_text(aes(label = confirmed), nudge_y = 0.0001, hjust = "left", size = 5) +
    #scale_y_continuous(breaks = seq(0, 0.01, by = 0.01)) +
    coord_flip() +
    labs(x = NULL,
         y = "Proportion confirmed relative to population size (text label: absolute value)",
         subtitle = "Selection of country inactive")
})
```

### World - top 40 confirmed

```{r}
renderPlot({
  DATA %>% 
    filter(date == max(date)) %>%
    arrange(desc(confirmed)) %>% 
    slice(1:40) %>% 
    ggplot(aes(reorder(country, confirmed), confirmed)) +
    theme_bw(base_size = 14) +
    geom_pointrange(aes(ymin = 0, ymax = confirmed), lwd = 1) +   
    #geom_text(aes(label = confirmed), nudge_y = 2, hjust = "left", size = 5) +
    coord_flip() +
    scale_y_log10() +
    labs(title = "Number of confirmed cases",
         subtitle = "Selection of country inactive",
         x = NULL,
         y = NULL)
})
```


Deaths
=====================================

Column {.tabset}
-------------------------------------

### Testing (Number of deaths per million by day after 1 death per million)

```{r}
uiOutput("high_deaths")
```

### Europe only

```{r}
renderPlot({
  DATA %>% 
    filter(date == max(date),
           country %in% countries.europa) %>% 
    mutate(p = deaths / confirmed) %>% 
    ggplot(aes(reorder(country, p), p)) +
    theme_bw(base_size = 14) +
    geom_pointrange(aes(ymin = 0, ymax = p), lwd = 1) +
    geom_text(aes(label = deaths), nudge_y = 0.001, hjust = "left", size = 5) +
    scale_y_continuous(breaks = seq(0, 0.2, by = 0.01)) +
    coord_flip() +
    labs(x = NULL,
         y = "Proportion dead relative to numbers confirmed (text label: absolute value)",
         subtitle = "Selection of country inactive")
})
```

### World

```{r}
renderPlot({
  DATA %>% 
    filter(date == max(date),
           deaths >= 5) %>% 
    mutate(p = deaths / confirmed) %>% 
    ggplot(aes(reorder(country, p), p)) +
    theme_bw(base_size = 14) +
    geom_pointrange(aes(ymin = 0, ymax = p), lwd = 1) +   
    geom_text(aes(label = deaths), nudge_y = 0.001, hjust = "left", size = 5) +
    coord_flip() +
    scale_y_continuous(breaks = seq(0, 20, by = 0.01)) +
    labs(x = NULL,
         y = "Proportion dead relative to numbers confirmed (text label: absolute value)",
         subtitle = "Selection of country inactive - Showing countries with 5 or more deaths")
})
```


Info
=====================================  

Confirmed graph:

* Grey lines: All european countries
* Yellow line: Median of european countries
* Green line: Confirmed profile of country in panel
* Red line: Death profile of country in panel (Note: deaths higher than 1 per million only shown)

Source:

* [Center for Systems Science and Engineering (CSSE) at Johns Hopkins University (JHU)
](https://gisanddata.maps.arcgis.com/apps/opsdashboard/index.html#/bda7594740fd40299423467b48e9ecf6)
* [Our world in data](https://ourworldindata.org/covid-testing)


SMALL PRINT: The data are heavily biased, hence making inference is not easy.

The code: [github](https://gist.github.com/einarhjorleifsson/cf22ddabedcc286c05965a080ef1cb08)
